\titlea\titleb\titlec\titled\titlee\titlef\titleg

### 2016-09-01 oopsla-16/overview.tex
oopsla: clean up/remove outdated comments in main tex

\revisionx{
{\bf Philosophy of bugs.}
% two camps: all races are bugs; citations:
% and, races must be classified; citations: replay analysis, portend,
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock)\footnote{
C++ declares any race between two non-atomic locations to be undefined behaviour \cite{cpp-foundation}.
From a C++ perspective, we assume all concurrent accesses are implemented by {\tt std::atomic} loads and stores.}.
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}

>>>>

\revisionx{
{\bf Philosophy of bugs.}
While there is a vast body of work on how to detect data races to begin with,
judging data races once found is a matter of philosophical debate unto itself.
Some recent tools classify races depending on how they impact program behaviour \cite{recordreplaydrs,portend}, overlooking {\em benign} races in search of more dangerous ones.
%In particular,
\cite{racerx} acknowledges that a program may have too many races for a user to worry about,
so bug reports must be prioritized by severity of effects. %how likely they may cause failures or security vulnerabilities.
%
However, other prior work argues that data races are always bugs \cite{miscompile-benign,data-races-are-evil},
largely due to the possibility of compiler or hardware reordering of racy accesses.
%
We take the former camp:
we consider a data race a bug only when it results in a visible failure state (e.g., crash or deadlock)\footnote{
C++ declares any race between two non-atomic locations to be undefined behaviour \cite{cpp-foundation}.
From a C++ perspective, we assume all concurrent accesses are implemented by {\tt std::atomic} loads and stores.}.
We bypass concerns of compiler reordering by checking programs at the executable level;
for a study of hardware reordering in the context of DPOR, we refer the reader to \cite{tsopso}.
}

### 2016-09-01 oopsla-16/paper.tex
oopsla: ack PDL consortium, make graph a bit bigger
oopsla: last revisions :O
oopsla: clean up/remove outdated comments in main tex

\newcommand\numthrlibs{79}
\newcommand\numpintoses{78} % 'di' is basecode. durr
\newcommand\numstudence{157} % total pintoses plus p2s

@@

% presubmit revisions
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}
% 2nd round revisions
%\newcommand\revisionx[1]{\hilight{purple}{#1}\xspace}
\newcommand\revisionx[1]{#1}
% cam ready revisions
\newcommand\revisionz[1]{\hilight{purple}{#1}\xspace}

\title{Stateless Model Checking with Data-Race Preemption Points}

\authorinfo{Ben Blum}{Carnegie Mellon University, USA}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University, USA}{garth@cs.cmu.edu}
%\authorinfo{\vspace{24pt}}{}{}

@@

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and automatically finds state spaces that are appropriately sized to complete in a given CPU budget.
%and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.
Between data-race bugs and verifications, \quicksand~is shown to provide the best of both worlds.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.

@@

%\section{Acknowledgements}

%\appendix
%\section{Appendix Title}
%
%This is the text of the appendix, if you need one.

@@

Many thanks to Vaishaal Shankar, Haryadi Gunawi, and David A. Eckhardt for generously providing student implementations from Berkeley's, U. of Chicago's, and CMU's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics.
Thanks to
Ji\v{r}\'{i} \v{S}im\v{s}a, Joshua Wise, Michael J. Sullivan, Carlo Angiuli, Jean Yang, Brandon Lucia, Haryadi Gunawi, John Wilkes, and
the anonymous OOPSLA reviewers for their helpful comments.
This work was supported in part by
the U.S. Army Research Office under grant number W911NF0910273.

>>>>

\newcommand\numthrlibs{79}
\newcommand\numpintoses{78}
\newcommand\numstudence{157} % total pintoses plus p2s

@@

% presubmit revisions
\newcommand\revision[1]{#1}
\definecolor{purple}{RGB}{192,0,192}
%\newcommand\revision[1]{\hilight{purple}{#1}}
% 2nd round revisions
%\newcommand\revisionx[1]{\hilight{purple}{#1}\xspace}
\newcommand\revisionx[1]{#1}
% cam ready revisions
%\newcommand\revisionz[1]{\hilight{purple}{#1}\xspace}
\newcommand\revisionz[1]{#1\xspace}

% obfuscate title just a little bit to avoid the billion inevitable google searches
% i don't mind if really dedicated people find their way here, but not everyone please
\newcommand\titlec{l Check}
\newcommand\titleg{ion Points}
\newcommand\titleb{less Mode}
\newcommand\titlee{h Data-R}
\newcommand\titled{ing wit}
\newcommand\titlef{ace Preempt}
\newcommand\titlea{State}

\title{\titlea\titleb\titlec\titled\titlee\titlef\titleg}

\authorinfo{Ben Blum}{Carnegie Mellon University, USA}{bblum@cs.cmu.edu}
\authorinfo{Garth Gibson}{Carnegie Mellon University, USA}{garth@cs.cmu.edu}

@@

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and automatically finds state spaces that are appropriately sized to complete in a given CPU budget.

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.
Between data-race bugs and verifications, \quicksand~is shown to provide the best of both worlds.
By using \numstudence~student operating system implementation projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.

@@

@@

Many thanks to Vaishaal Shankar, Haryadi Gunawi, and David A. Eckhardt for generously providing student implementations from Berkeley's, U. of Chicago's, and CMU's OS classes respectively.
Thanks to Wind River for the use of their simulator \simics.
Thanks to
Ji\v{r}\'{i} \v{S}im\v{s}a, Joshua Wise, Michael J. Sullivan, Carlo Angiuli, Jean Yang, Brandon Lucia, Haryadi Gunawi, John Wilkes, and
the anonymous OOPSLA reviewers for their helpful comments.
%
We also thank the member companies of the PDL Consortium (Broadcom,
Citadel, EMC, Facebook, Google, Hewlett-Packard Labs, Hitachi, Intel,
Microsoft Research, MongoDB, NetApp, Oracle, Samsung America, Seagate,
Tintri, Two Sigma, Uber, Veritas and Western Digital).
%
This work was supported in part by
the U.S. Army Research Office under grant number W911NF0910273.

### 2016-09-01 oopsla-16/related.tex
oopsla: clean up/remove outdated comments in main tex

We build upon many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%, and Eraser, the original data race detector \cite{eraser}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related tools by their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence, they will miss any bugs that require interleaving threads at instruction granularity during a data race.
%CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.
\revision{\cite{chess} discusses the ability of CHESS to add PPs using a single-pass data-race analysis,
% FIXME... idk, "discusses (without explicit evaluation)..., but, as our evaluation shows..."?
but does not evaluate either the increase in bug-finding capacity or the soundness properties.
Our convergence theorem (\sect{\ref{sec:soundness}}) is a natural extension of \cite{chess-icb}'s Theorem 3,
which alone provides soundness only for race-free programs.}

{\bf Message-passing.} Other stateless model checkers, such as SAMC \cite{samc}, MaceMC \cite{macemc}, MoDist \cite{modist}, ETA \cite{dbug-retreat}, and Concuerror \cite{optimal-dpor},
limit thread communication to a message-passing API to more effectively test distributed systems.
This eliminates the need for data-race analysis, but restricts the class of programs that can be tested.
%Nevertheless, Iterative Deepening is applicable to these tools.

@@

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT \revisionx{targets} tests with impossibly large state spaces,
\revisionx{eschewing DPOR's depth-first search model to
%ensure it can
instead sample broad cross-sections of large state spaces.
However, it proposes no alternate reduction algorithm to make up for its incompatibility with DPOR,
so is unsuitable for verification of medium-sized tests.}
%while for more approachable tests,
%its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
%
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB
%is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and can incorporate data-race PPs only when statically coded in advance}.
\revisionx{Our results in \sect{\ref{sec:eval-sssmc}} outperform both such configurations of ICB.}
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.

@@

\revisionx{{\bf Other domains.}}
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} \revision{finds data races in}~kernel code.
IFRit \cite{ifrit}
improves the performance of HB using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

@@

% \subsection{Other Concurrency Testing Approaches}
%
% blah blah pldi'15 symbiosis DSP

% Note that BPOR paper claims that ICB(3+) repeats LOADS of work, and that makes it ok for landslide-ID to repeat work.

% IDK if i should mention it, but OOPSLA 2015, protocol based verification of MPI concurrency paper. Different verification approach entirely; doesn't suffer exponential explosion but limited to programs with no shared state and MPI communication only

% Probably NOT worth a mention: OOPSLA 2015, stateless model checking of event driven applications. Turning timer-driven model on its head and checking single-threaded, but asynch-event-driven programs (i.e. device-like signal handlers)

>>>>

We build upon many established model-checking techniques, dating back
% of course
to Verisoft, the original C model checker \cite{verisoft}.
%Our model checker
%\landslide~\cite{landslide} itself implements many techniques from prior work (\sect{\ref{sec:landslide}}).
%itself implements DPOR \cite{dpor},
%state space estimation \cite{estimation},
%and data-race detection \cite{eraser}.
We compare related tools by their treatment of shared-memory thread communication.

{\bf Synchronization events only.} CHESS \cite{chess} and dBug \cite{dbug-ssv} instrument the thread library API, and can preempt programs only during calls to this API.
Hence, they will miss any bugs that require interleaving threads at instruction granularity during a data race.
%CHESS provides a data-race analysis to report any such violations of its concurrency model to the user, but does not incorporate data-race candidates as PPs in future tests.
\revision{\cite{chess} discusses the ability of CHESS to add PPs using a single-pass data-race analysis,
but does not evaluate either the increase in bug-finding capacity or the soundness properties.
Our convergence theorem (\sect{\ref{sec:soundness}}) is a natural extension of \cite{chess-icb}'s Theorem 3,
which alone provides soundness only for race-free programs.}

{\bf Message-passing.} Other stateless model checkers, such as SAMC \cite{samc}, MaceMC \cite{macemc}, MoDist \cite{modist}, ETA \cite{dbug-retreat}, and Concuerror \cite{optimal-dpor},
limit thread communication to a message-passing API to more effectively test distributed systems.
This eliminates the need for data-race analysis, but restricts the class of programs that can be tested.
%Nevertheless, Iterative Deepening is applicable to these tools.
% IDK if i should mention it, but OOPSLA 2015, protocol based verification of MPI concurrency paper. Different verification approach entirely; doesn't suffer exponential explosion but limited to programs with no shared state and MPI communication only

@@

\revision{{\bf Restricting preemptions.}
Preemption Sealing \cite{sealing} pre\-sents a mechanism similar to the {\tt within\_function} command (\sect{\ref{sec:landslide}}) for users to manually restrict preemptions.
It demonstrates the need to consider subsets of PPs,
as well as developers' willingness to limit a test case's scope so the resulting state space may be fully verified (\sect{\ref{sec:totalverif}}).
%It justifies our assumption that expert
%programmers can customize small test cases which, when fully verified,
%provide a correctness assurance much stronger than the assurance from
%a partially-complete exploration of a huge state space.
Probabilistic Concurrency Testing (PCT) \cite{randomized-scheduler}
is a randomized algorithm that can quantify the probability of uncovering bugs.
PCT \revisionx{targets} tests with impossibly large state spaces,
\revisionx{eschewing DPOR's depth-first search model to
%ensure it can
instead sample broad cross-sections of large state spaces.
However, it proposes no alternate reduction algorithm to make up for its incompatibility with DPOR,
so is unsuitable for verification of medium-sized tests.}
%while for more approachable tests,
%its incompatibility with DPOR would significantly delay completion.
Future work could use ETAs to heuristically switch between DPOR and PCT}.
%
Finally, ICB \cite{chess-icb} is most similar to our work,
as both approaches provide a partial verification
%on some subset of interleavings
when full completion is intractable (\sect{\ref{sec:future}}).
However, ICB
%is limited to a fixed set of PPs,
\revision{cannot estimate remaining time to total verification,
and can incorporate data-race PPs only when statically coded in advance}.
\revisionx{Our results in \sect{\ref{sec:eval-sssmc}} outperform both such configurations of ICB.}
%and to our knowledge no algorithm has been proposed to dynamically add data-race PPs during a test with ICB.
% Note that BPOR paper claims that ICB(3+) repeats LOADS of work, and that makes it ok for landslide-ID to repeat work.

@@

\revisionx{{\bf Other domains.}}
\revision{Several tools \cite{droidracer,cafa,unnamed-android} have recently emerged to target Android applications},
%DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%%event-driven
%Android applications,
using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} \revision{finds data races in}~kernel code.
IFRit \cite{ifrit}
improves the performance of HB using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).

@@
